---
title: "Challenge 8"
subtitle: "Out of an abundance of caution . . . "
author: Kyeri Kim
output:
  html_document:
    df_print: paged
---
# Abundance Plot Challenge

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, include=TRUE)
```

## Background
The goal of these challenges are to give you experience with absolute abundance plots, as well as with several features of phyloseq. For several technical reasons, the absolute abundance of amplicon sequencing reads is not biologically meaningful.  However, review of absolute abundance is important for quality control.

## Resources
These challenges are based on the [phyloseq plot_bar tutorial](https://joey711.github.io/phyloseq/plot_bar-examples.html), which will give you starting points for most of the challenges.  

## Data
For all of these challenges you will be using a lemur data phyloseq RDS that is available at `/data/tutorial_data/lemur_subset_phyloseq.rds`.

## Completing and Submitting
You must fill in the chunks below as indicated, but you are free to add other chunks too.  To submit this assignment for full credit you should *commit* and *push*:

1. This file (`challenge8_assignment.Rmd`)
2. The knited version of this file (`challenge8_assignment.html`)

These are the *ONLY* files you should include in your repo.  I strongly recommend that you make a `scratch` subdirectory in your home directory, then make subdirectories for each project in scratch, for example `~/scratch/challenge8`.  If you don't follow this advice, but instead put temporary files in your repo directory, you *must* be sure not to commit them to your repo, you will lose points if you do.



# Challenge Chunks
## Load Libraries
```{r}
library(readr)
library(phyloseq)
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(magrittr)
library(vegan)
```

## Define Paths
```{r}
poly.rds ="/sharedspace/polystyrene/demultiplexed/dada2/ibiemProject1_subset.rds"
```

## Check Phyloseq RDS
Load your phyloseq RDS and `print` it just to be sure things look sane
```{r}
poly.ps = read_rds(poly.rds)
print(poly.ps)
```

```{r}
# save sample data as a data frame
df <- sample_data(poly.ps)

week <- NULL
collectionID <- NULL
site <- NULL
group <- NULL

# create week and collectionID variables
for(i in 1:nrow(sample_data(poly.ps))){
  week[i] <- strsplit(sample_data(poly.ps)$Description, "_")[[i]][1]
  collectionID[i] <- as.numeric(strsplit(sample_data(poly.ps)$Description, "_")[[i]][2])
}

# create site variable
site[collectionID %in% 1:4] <- "LF"
site[collectionID %in% 5:8] <- "ER"
site[collectionID %in% 9:12] <- "AS"
site[collectionID %in% 13:16] <- "RC"

# create group variable 
group[mod(collectionID, 4) == 1] <- "PS"
group[mod(collectionID, 4) == 2] <- "S"
group[mod(collectionID, 4) == 3] <- "PSS"
group[mod(collectionID, 4) == 0] <- "C"

# add new variables to dataset
df$Week <- as.numeric(week)
df$Site <- site
df$Group <- group

# here is the new sample_data
df
df %>%
  select(Description, Week, Site, Group)

```

Add newly created sample_data file to phyloseq object
```{r}
poly.new.ps <- phyloseq(otu_table(poly.ps),
               df,
               tax_table(poly.ps))

```

Subset by LF
```{r}
ps_LF <- subset_samples(poly.new.ps, Site == "LF")
ps_ER <- subset_samples(poly.new.ps, Site == "ER")
ps_AS <- subset_samples(poly.new.ps, Site == "AS")
ps_RC <- subset_samples(poly.new.ps, Site == "RC")

```

```{r}
# landfill relative abundance plot
rel_abund_func <- function(ps, level, top){
  top20 <- names(sort(taxa_sums(ps), decreasing=TRUE))[1:top]
ps.top20 <- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
ps.top20<- prune_taxa(top20, ps.top20)

plot_bar(ps.top20, x="Week", fill= level) + 
  geom_bar(stat = "identity", position = "stack", size=0) +
  facet_wrap(~Group, scales="free_x") 
}

```

```{r}
rel_abund_func(ps_LF, "Family", 20)
rel_abund_func(ps_ER, "Family", 20)
rel_abund_func(ps_AS, "Family", 20)
rel_abund_func(ps_RC, "Family", 20)
```

Plot richness using Shannon
```{r}
plot_richness(poly.new.ps, x="Week", measures=c("Shannon"), color="Group") + 
  theme_bw() +
  facet_wrap(~Site)
```

Plot richness using Simpson
```{r}
# simpson
plot_richness(poly.new.ps, x="Week", measures=c("Simpson"), color="Group") + 
  theme_bw() +
  facet_wrap(~Site)
```

## Relative Abundance Barplot

```{r}
# top 20 taxa across all sites and conditions
top20 <- names(sort(taxa_sums(poly.new.ps), decreasing=TRUE))[1:20]
ps.top20_rel <- transform_sample_counts(poly.new.ps, function(OTU) OTU/sum(OTU))
ps.top20 <- prune_taxa(top20, ps.top20_rel)

# check sample sums all add to 1
sample_sums(ps.top20_rel)

# plot ALL relative abundance in one plot
plot_bar(ps.top20, fill = "Family") +
  labs(y = "Relative Abundance")

# facet by site and group
plot_bar(ps.top20, x="Week", fill="Family") + 
  geom_bar(stat = "identity", position = "stack", size=0) +
  facet_wrap(~Site + Group, scales="free_x") 

```


Plot relative abundance by genus level
```{r}
rel_abund_func_level(ps_LF, "Genus", 20)
rel_abund_func_level(ps_ER, "Genus", 20)
rel_abund_func_level(ps_AS, "Genus", 20)
rel_abund_func_level(ps_RC, "Genus", 20)
```



```{r}
# how to compare increasingness
plot_bar(ps.top20, x="Description", fill="Genus") +
    geom_point(size = 2) +
  geom_line() +
  facet_wrap(~Site, scales="free_x") 
```

```{r}
mdf <-  psmelt(ps.top20)
p <- ggplot(mdf, aes_string(x = "Week", y = "Abundance", color = "Genus")) +
  geom_point(size = 2) +
  geom_line() +
  facet_wrap(~Site, scales="free_x")
p
```

## Permanova

Don't need to prune because all samples have large counts
```{r}
sample_sums(poly.new.ps) %>% sort
```
```{r}
min_count = 3
min_sample = 2

prune.vec = filter_taxa(poly.new.ps, 
                       function(x) sum(x >= min_count) >= min_sample)
sum(prune.vec)
```

Transform to even sampling depth
```{r}
ps_even = transform_sample_counts(poly.new.ps, function(x) 1E6 * x/sum(x))

ps.st_prune.even = prune_taxa(prune.vec, ps_even)
ntaxa(ps.st_prune.even)
```
```{r}
# plot_bar(ps.st_prune.even)
```

We want to see if there are differences in microbial communities by sites and groups. A permanova analysis allows us to quantify this.
```{r}
ps.st_prune.even %>%
  sample_data %>%
  as("data.frame") ->
  ps.st_prune.even.metadata
```

```{r}
ps.st_prune.even %>%
  distance(method="bray") ->
  ps.st_prune.even.bray

```

Permanova by Site and Group
```{r}
adonis(ps.st_prune.even.bray ~ Site + Group,
       data = ps.st_prune.even.metadata) ->
  site_adonis

print(site_adonis)
```

Permanova by group


# Session Info
Always print `sessionInfo` for reproducibility!
```{r}
sessionInfo()
```

