---
author: Kyeri Kim
output:
  html_document:
    df_print: paged
---
# Abundance Plot Challenge

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, include=TRUE)
```

## Background
The goal of these challenges are to give you experience with absolute abundance plots, as well as with several features of phyloseq. For several technical reasons, the absolute abundance of amplicon sequencing reads is not biologically meaningful.  However, review of absolute abundance is important for quality control.

## Resources
These challenges are based on the [phyloseq plot_bar tutorial](https://joey711.github.io/phyloseq/plot_bar-examples.html), which will give you starting points for most of the challenges.  

## Data
For all of these challenges you will be using a lemur data phyloseq RDS that is available at `/data/tutorial_data/lemur_subset_phyloseq.rds`.

## Completing and Submitting
You must fill in the chunks below as indicated, but you are free to add other chunks too.  To submit this assignment for full credit you should *commit* and *push*:

1. This file (`challenge8_assignment.Rmd`)
2. The knited version of this file (`challenge8_assignment.html`)

These are the *ONLY* files you should include in your repo.  I strongly recommend that you make a `scratch` subdirectory in your home directory, then make subdirectories for each project in scratch, for example `~/scratch/challenge8`.  If you don't follow this advice, but instead put temporary files in your repo directory, you *must* be sure not to commit them to your repo, you will lose points if you do.



# Challenge Chunks
## Load Libraries
```{r}
library(readr)
library(phyloseq)
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(magrittr)
library(vegan)
```

## Define Paths
```{r}
poly.rds ="/sharedspace/polystyrene/demultiplexed5/dada2/ibiemProject1_subset.rds"
#poly.rds ="/home/guest/project1_ibiem/demultiplexed3/dada2/ibiemProject1_subset.rds"
```

## Check Phyloseq RDS
Load your phyloseq RDS and `print` it just to be sure things look sane
```{r}
poly.ps = read_rds(poly.rds)
print(poly.ps)
```
## Subset by location
```{r}
ps_LF <- subset_samples(poly.ps, Location=="LF")
ps_ER <- subset_samples(poly.ps, Location=="ER")
ps_AS <- subset_samples(poly.ps, Location=="AS")
ps_RC <- subset_samples(poly.ps, Location=="RC")
ps_LF
ps_ER
ps_AS
ps_RC
```
# landfill relative abundance plot
```{r}
# rel_abund_func <- function(ps, level, top){
#   top20 <- names(sort(taxa_sums(ps), decreasing=TRUE))[1:top]
# ps.top20 <- transform_sample_counts(ps, function(OTU) OTU/sum(OTU))
# ps.top20<- prune_taxa(top20, ps.top20)
# 
# plot_bar(ps.top20, x="Week", fill= level) + 
#   geom_bar(stat = "identity", position = "stack", size=0) +
#   facet_wrap(~Condition)#, scales="free_x") 
# }

```

```{r}
# rel_abund_func(ps_LF, "Family", 20)
# rel_abund_func(ps_ER, "Family", 20) +
#   theme(legend.text  = element_text(size = 8),
#         legend.key.size = unit(.6, "cm"))
# rel_abund_func(ps_AS, "Family", 20)
# rel_abund_func(ps_RC, "Family", 20)
```

## Transform OTU -> relative OTU
```{r}
rel_ps_LF <-transform_sample_counts(ps_LF, function(OTU) OTU/sum(OTU))
rel_ps_ER <-transform_sample_counts(ps_ER, function(OTU) OTU/sum(OTU))
rel_ps_AS <-transform_sample_counts(ps_AS, function(OTU) OTU/sum(OTU))
rel_ps_RC <-transform_sample_counts(ps_RC, function(OTU) OTU/sum(OTU))

as.data.frame(otu_table(subset_samples(rel_ps_LF, Condition =="PS"))[ , 1])
```
## Plot a specific taxa
```{r}
#taxa table has the information of each OTU
tax_table(rel_ps_LF)[1,]

Pseudo <- subset_taxa(rel_ps_LF, Order=="Pseudomonadales")
plot_bar(Pseudo, x="Week", fill="Family") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+
  facet_wrap(~Condition, nrow=1) +
  ggtitle("Pseudomonadales in LF")

plot_bar(Pseudo, x="Week", fill="Genus") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+
  facet_wrap(~Condition, nrow=1) +
  ggtitle("Pseudomonadales in LF")
```

# sweep func test: substract row-wise with a vector
```{r}
# 
# standardweek <-1
# otu_LF_Conditioned <- sweep(otu_table(subset_samples(rel_ps_LF, Condition=="PS")),2,as.double(otu_table(subset_samples(rel_ps_LF, Week==standardweek & Condition=="PS"))))
# 
# #number of seqences
# threshold = 0.01
# sum(otu_LF_Conditioned[2,]>threshold)
# 
# #sort seq. names of "increased" OTU (Week12 - Week1)
# a<-otu_LF_Conditioned[2,]
# targets <- names(a[otu_LF_Conditioned[2,]>threshold])
# 
# target.ps <- prune_taxa(targets, subset_samples(rel_ps_LF, Condition=="PS"))
# plot_bar(target.ps, x="Week", fill="Family") + 
#   geom_bar(stat = "identity", position = "stack", size=0) 

```
```{r}
standardweek <-1
threshold = 0.01

for (cond in c("PS", "PSS","S", "Ctrl")) {
#sweep: substract row-wise with a vector
otu_LF_Conditioned <- sweep(otu_table(subset_samples(rel_ps_LF, Condition==cond)),2,as.double(otu_table(subset_samples(rel_ps_LF, Week==standardweek & Condition==cond))))

#number of seqences
sum(otu_LF_Conditioned[2,]>threshold)

#sort seq. names of "increased" OTU (Week12 - Week1)
a<-otu_LF_Conditioned[2,]
targets <- names(a[a>threshold])

target.ps <- prune_taxa(targets, subset_samples(rel_ps_LF, Condition==cond))
print( plot_bar(target.ps, x="Week", fill="Family") +
  geom_bar(stat = "identity", position = "stack", size=0)
)
}
```



## weed N - week 1 rel. abundance in for loop
```{r}
#"LF", "ER", "AS","RC"   c(rel_ps_LF, rel_ps_ER, rel_ps_AS, rel_ps_RC)
standardweek <-1
threshold = 0
i = 0
for (loc_ps in c(rel_ps_LF, rel_ps_ER, rel_ps_AS, rel_ps_RC)){
  i = i+1
  loc <- c("LF", "ER", "AS","RC" )[i]
    for (cond in c("PS", "PSS","S", "Ctrl")) {
    #sweep: substract row-wise with a vector
    otu_Conditioned <- sweep(otu_table(subset_samples(loc_ps, Condition==cond)), 2, as.double(otu_table(subset_samples(loc_ps, Week==standardweek & Condition==cond))))
    
    #number of seqences
    # sum(otu_Conditioned[2,]>threshold)
    
    #sort seq. names of "increased" OTU (Week12 - Week1) : 
    # Week 12 = [2, ] Week7 = [8,]
    a<-otu_Conditioned[2,]
    b<-otu_Conditioned[8,]
    
    #assign target variable name by location and condition for further use: names will be used
    # assign(paste(loc, cond, "Degraders", sep = "_"), c(names(a[otu_Conditioned[2,]>threshold])))
    targets_a <- names(a[a>threshold])
    targets_b <- names(b[b>threshold])
    
    target_a.ps <- prune_taxa(targets_a, subset_samples(loc_ps, Condition==cond))
    target_b.ps <- prune_taxa(targets_b, subset_samples(loc_ps, Condition==cond))
    assign(paste(loc, cond, "Degraders12", sep = "_"), targets_a)
    assign(paste(loc, cond, "Degraders7", sep = "_"), targets_b)
    
    # plot12 <- plot_bar(target_a.ps, x="Week", fill="Family") +  geom_bar(stat = "identity", position = "stack", size=0)
    # plot7  <- plot_bar(target_b.ps, x="Week", fill="Family") +  geom_bar(stat = "identity", position = "stack", size=0)
    # assign(paste(loc, cond, "Plot12", sep = "_"), plot12)
    # assign(paste(loc, cond, "Plot7", sep = "_"), plot7)
    }
}
```
## Check the assinged plot name/otu
```{r}
head(LF_PSS_Degraders12)

head(LF_PSS_Degraders7)
# #"LF", "ER", "AS","RC"   c(rel_ps_LF, rel_ps_ER, rel_ps_AS, rel_ps_RC)
# LF_PS_Plot
# LF_PSS_Plot
# LF_S_Plot
# LF_Ctrl_Plot
# rel_ps_LF[,1]
```

## Generate VennDiagram Week12=Week1
```{r}
# install.packages("VennDiagram")
library(VennDiagram)

# Chart
venn.diagram(
  x = list(LF_PSS_Degraders12, LF_PS_Degraders12, LF_S_Degraders12, LF_Ctrl_Degraders12),
  category.names = c("PSS" , "PS" , "S", "Ctrl"),
  main = "LF",
  filename = 'KYERI_Img/w12/#LF_venn_diagramm_th0.png',
  output=TRUE
)
venn.diagram(
  x = list(AS_PSS_Degraders12, AS_PS_Degraders12, AS_S_Degraders12, AS_Ctrl_Degraders12),
  category.names = c("PSS" , "PS" , "S", "Ctrl"),
  main = "AS",
  filename = 'KYERI_Img/w12/#AS_venn_diagramm_th0.png',
  output=TRUE
)
venn.diagram(
  x = list(RC_PSS_Degraders12, RC_PS_Degraders12, RC_S_Degraders12, RC_Ctrl_Degraders12),
  category.names = c("PSS" , "PS" , "S", "Ctrl"),
  main = "RC",
  filename = 'KYERI_Img/w12/#RC_venn_diagramm_th0.png',
  output=TRUE
)
venn.diagram(
  x = list(ER_PSS_Degraders12, ER_PS_Degraders12, ER_S_Degraders12, ER_Ctrl_Degraders12),
  category.names = c("PSS" , "PS" , "S", "Ctrl"),
  main = "ER",
  filename = 'KYERI_Img/w12/#ER_venn_diagramm_th0.png',
  output=TRUE
)
```

## Generate VennDiagram Week12=Week1
```{r}
# install.packages("VennDiagram")
library(VennDiagram)

# Chart
venn.diagram(
  x = list(LF_PSS_Degraders7, LF_PS_Degraders7, LF_S_Degraders7, LF_Ctrl_Degraders7),
  category.names = c("PSS" , "PS" , "S", "Ctrl"),
  main = "LF",
  filename = 'KYERI_Img/w7/#LF_venn_diagramm_th0.png',
  output=TRUE
)
venn.diagram(
  x = list(AS_PSS_Degraders7, AS_PS_Degraders7, AS_S_Degraders7, AS_Ctrl_Degraders7),
  category.names = c("PSS" , "PS" , "S", "Ctrl"),
  main = "AS",
  filename = 'KYERI_Img/w7/#AS_venn_diagramm_th0.png',
  output=TRUE
)
venn.diagram(
  x = list(RC_PSS_Degraders7, RC_PS_Degraders7, RC_S_Degraders7, RC_Ctrl_Degraders7),
  category.names = c("PSS" , "PS" , "S", "Ctrl"),
  main = "RC",
  filename = 'KYERI_Img/w7/#RC_venn_diagramm_th0.png',
  output=TRUE
)
venn.diagram(
  x = list(ER_PSS_Degraders7, ER_PS_Degraders7, ER_S_Degraders7, ER_Ctrl_Degraders7),
  category.names = c("PSS" , "PS" , "S", "Ctrl"),
  main = "ER",
  filename = 'KYERI_Img/w7/#ER_venn_diagramm_th0.png',
  output=TRUE
)
```


## Intersect function (to find commons) & setdiff fuction (to get rid of commons in groups)

```{r}
intS <-c(intersect(c(1:10), c(5:15)))
intS
setdiff(intS, c(7:10))
```
```{r}
# intersect_taxa(LF_PSS_Degraders, LF_PS_Degraders)
intersect(LF_PSS_Degraders12, LF_PS_Degraders12)

```






## Landfill 
```{r}
#week 12
aa <- Reduce(intersect, list(LF_PSS_Degraders12, LF_PS_Degraders12, LF_S_Degraders12)) # 3 common - Ctrl#
bb <- setdiff(aa, LF_Ctrl_Degraders12)
length(bb) 
Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
tax_table(Degrader.ps)# the only one hit - no assigned genus
plot_bar(Degrader.ps, x="Week", fill="Family") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
ggsave("KYERI_Img/w12/#LF_PS+PSS+S-C.png")

#week7
aa <- Reduce(intersect, list(LF_PSS_Degraders7, LF_PS_Degraders7, LF_S_Degraders7)) # 3 common - Ctrl#
bb <- setdiff(aa, LF_Ctrl_Degraders7)

Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
Degrader.ps
plot_bar(Degrader.ps, x="Week", fill="Genus") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
ggsave("KYERI_Img/w7/#LF_PS+PSS+S-C.png")


# PS & PSS but not in S &Ctrl
aa <- Reduce(intersect, list(LF_PSS_Degraders7, LF_PS_Degraders7)) # 3 common - Ctrl#
bb <- setdiff(aa, LF_Ctrl_Degraders7)
bb <- setdiff(bb, LF_S_Degraders7)
Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
Degrader.ps
plot_bar(Degrader.ps, x="Week", fill="Genus") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
```
## ER
```{r}
#week 12
aa <- Reduce(intersect, list(ER_PSS_Degraders12, ER_PS_Degraders12, ER_S_Degraders12)) # 3 common - Ctrl#
bb <- setdiff(aa, ER_Ctrl_Degraders12)
length(bb) 
Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
tax_table(Degrader.ps)# the only one hit - no assigned famiy
plot_bar(Degrader.ps, x="Week", fill="Order") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
ggsave("KYERI_Img/w12/#ER_PS+PSS+S-C.png")

#week7
aa <- Reduce(intersect, list(ER_PSS_Degraders7, ER_PS_Degraders7, ER_S_Degraders7)) # 3 common - Ctrl#
bb <- setdiff(aa, ER_Ctrl_Degraders7)

Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
Degrader.ps
plot_bar(Degrader.ps, x="Week", fill="Genus") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
ggsave("KYERI_Img/w7/#ER_PS+PSS+S-C.png")

# bb <- setdiff(ER_PS_Degraders7, ER_Ctrl_Degraders7)
# Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
# plot_bar(Degrader.ps, x="Week", fill="Family") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
```
## AS
```{r}
#week 12
aa <- Reduce(intersect, list(AS_PSS_Degraders12, AS_PS_Degraders12, AS_S_Degraders12)) # 3 common - Ctrl#
bb <- setdiff(aa, AS_Ctrl_Degraders12)
length(bb) 
Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
tax_table(Degrader.ps)# the only one hit - no assigned famiy
plot_bar(Degrader.ps, x="Week", fill="Order") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
ggsave("KYERI_Img/w12/#AS_PS+PSS+S-C.png")

#week7
aa <- Reduce(intersect, list(AS_PSS_Degraders7, AS_PS_Degraders7, AS_S_Degraders7)) # 3 common - Ctrl#
bb <- setdiff(aa, AS_Ctrl_Degraders7)

Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
Degrader.ps
plot_bar(Degrader.ps, x="Week", fill="Genus") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
ggsave("KYERI_Img/w7/#AS_PS+PSS+S-C.png")

bb <- setdiff(AS_PS_Degraders7, AS_Ctrl_Degraders7)
Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
plot_bar(Degrader.ps, x="Week", fill="Family") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
```
## RC
```{r}
#week 12
aa <- Reduce(intersect, list(RC_PSS_Degraders12, RC_PS_Degraders12, RC_S_Degraders12)) # 3 common - Ctrl#
bb <- setdiff(aa, RC_Ctrl_Degraders12)
length(bb) 
Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
tax_table(Degrader.ps)# the only one hit - no assigned famiy
plot_bar(Degrader.ps, x="Week", fill="Order") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
ggsave("KYERI_Img/w12/#RC_PS+PSS+S-C.png")

#week7
aa <- Reduce(intersect, list(RC_PSS_Degraders7, RC_PS_Degraders7, RC_S_Degraders7)) # 3 common - Ctrl#
bb <- setdiff(aa, RC_Ctrl_Degraders7)

Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
Degrader.ps
plot_bar(Degrader.ps, x="Week", fill="Genus") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
ggsave("KYERI_Img/w7/#RC_PS+PSS+S-C.png")

# bb <- setdiff(RC_PS_Degraders7, RC_Ctrl_Degraders7)
# Degrader.ps <- prune_taxa(bb, subset_samples(rel_ps_LF))
# plot_bar(Degrader.ps, x="Week", fill="Family") +  geom_bar(stat = "identity", position = "stack", size=0)+ scale_x_discrete(limits=c(1:7,12))+theme_bw()
```
https://www.ebi.ac.uk/biomodels-main/path2models?about=205







# Session Info
Always print `sessionInfo` for reproducibility!
```{r}
sessionInfo()
```

