---
title: 'Linear Model: Polystyrene Analysis'
author: "Kara McCormack"
output:
  pdf_document: default
  html_notebook: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, include=TRUE)
```

## Load Libraries
```{r}
library(readr)
library(phyloseq)
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyr)
library(magrittr)
library(vegan)
```

## Goals: 
1. Subset by location
2. Subset by condition
3. Conduct a linear regression with the following variables:
  * x = Week
  * y = OTU relative abundance
4. Order by $R^2$ values, want to use those with high $R^2$ value
5. Order by coefficient value on week, want to use those with higher coefficient
  * this mean that time is significant for relative abundance for that OTU
  * this means that that OTU might be a polystyrene degrader
  

## Define Paths
```{r}
poly.rds ="/sharedspace/polystyrene/demultiplexed5/dada2/ibiemProject1_subset.rds"
```

## Check Phyloseq RDS
Load your phyloseq RDS and `print` it just to be sure things look sane
```{r}
poly.ps = read_rds(poly.rds)
print(poly.ps)
```

## Rename Week variable
```{r}
# calculate relative abundance
rel <-transform_sample_counts(poly.ps, function(OTU) OTU/sum(OTU))

# turn the otu table into a data.frame
otu <- as.data.frame(otu_table(rel))
head(otu)

# grab first and last OTU's names for referencing columns in pivot_longer
first_otu <- colnames(otu)[1]
last_otu <- colnames(otu)[ncol(otu)]

otu %>%
  pivot_longer(first_otu:last_otu, 
               names_to = "OTU", 
               values_to = "Relative Abundance") %>%
  head
```

## Subset samples
```{r}
ps_LF <- subset_samples(poly.ps, Location=="LF")
ps_ER <- subset_samples(poly.ps, Location=="ER")
ps_AS <- subset_samples(poly.ps, Location=="AS")
ps_RC <- subset_samples(poly.ps, Location=="RC")
```

## Calculate relative abundance
```{r}
rel_ps_LF <-transform_sample_counts(ps_LF, function(OTU) OTU/sum(OTU))
rel_ps_ER <-transform_sample_counts(ps_ER, function(OTU) OTU/sum(OTU))
rel_ps_AS <-transform_sample_counts(ps_AS, function(OTU) OTU/sum(OTU))
rel_ps_RC <-transform_sample_counts(ps_RC, function(OTU) OTU/sum(OTU))
```


## Subset by Condition
We break apart the Landfill phyloseq object into four categories:

* Control
* Polystyrene
* Styrene
* Polystyrene + Styrene

```{r}
LF_ctrl <- subset_samples(rel_ps_LF, Condition == "Ctrl")
LF_ps <- subset_samples(rel_ps_LF, Condition == "PS")
LF_s <- subset_samples(rel_ps_LF, Condition == "S")
LF_pss <- subset_samples(rel_ps_LF, Condition == "PSS")
```

### Begin threshold_function items

## Subset to those with nonzero values for week 7 and 12
We will filter out those with less than 0.1% relative abundance at weeks 7 and 12, because we want to look at those that have some substantial presence at those weeks, in order to even consider them as polystyrene degraders.
```{r}
# Grab OTUs from control and make into data frame
LF_df_ctrl <- as.data.frame(otu_table(LF_ctrl))
LF_df_ps <- as.data.frame(otu_table(LF_ps))
LF_df_s <- as.data.frame(otu_table(LF_s))
LF_df_pss <- as.data.frame(otu_table(LF_pss))

```

```{r}
# reorder data frame so that weeks are in correct order
LF_df_ctrl <- LF_df_ctrl[c(1, 3:8, 2),]
LF_df_ps <- LF_df_ps[c(1, 3:8, 2),]
LF_df_s <- LF_df_s[c(1, 3:8, 2),]
LF_df_pss <- LF_df_pss[c(1, 3:8, 2),]
```

```{r}
# add week as a column
Week <- c(1:7, 12)

# Landfill 
LF_df_ctrl <- LF_df_ctrl %>%
  mutate(Week = as.factor(Week)) %>%
  select(Week, everything())
```

```{r}
# subset into only those OTUs with greater than 0.1% at weeks 7 and 12
subset <- LF_df_ctrl %>%
  pivot_longer(-Week,
               names_to = "OTU", 
               values_to = "Relative Abundance") %>%
  filter(Week == 7 | Week == 12) %>% 
  filter(`Relative Abundance` > .001) 

dim(subset)

otu_names <- unique(subset$OTU)
length(otu_names)
```
So we have 383 total samples in the Landfill site where, in the control condition, the OTUs had greater than .1% abundance at both weeks 7 and 12. Since some of these exceeded the threshold at both weeks 7 and 12, we see that there are a total of 220 unique OTUs satisfying this requirement. 

Now, we need to perform linear regression on only those OTUs. 

First, we will subset the OTU table using only those OTUs satisfying the threshold requirement. 
```{r}
LF_df_ctrl_subset <- LF_df_ctrl[,colnames(LF_df_ctrl)%in%otu_names]
dim(LF_df_ctrl_subset)
```
Now, we will loop through these OTUs, performing linear regression on each relative abundance as a function of Week. 

```{r}
options(scipen = 999)
week <- c(1:7, 12)
result <- NULL
for(otu in 2:ncol(LF_df_ctrl_subset)){
  my_col <- as.vector(LF_df_ctrl_subset[,otu])
  result[otu] <- round(summary(lm(my_col ~ week))$r.squared, digits = 3)
}
result[order(result, decreasing = T)]
```
Now we list OTUs that had an R-squared above 0.5. 
```{r}
# count the number of OTUs with R-squared values above 0.5
num_above <- sum(result[order(result, decreasing = T)] >.5, na.rm = T)

# grab the indices of the OTUs with R-squared values above 0.5
ord <- order(result, decreasing = T)[1:num_above]

# subset the dataframe with only those OTUs. 
result_LF_ctrl <- LF_df_ctrl_subset[ , ord]

```


## Function to find those above relative abundance threshold
```{r}
# input: data frame of otu_table at site with specific condition
# output: dataframe of OTUs satisfying requirement for relative abundance threshold, how many satisfy the threshold

threshold_func <- function(df, threshold = .5){
  
}
```


## test run for one iteration
```{r}
# week <- c(1, 12, 2:7)
# my_otu_y <- as.vector(otu_table(ps_LF_ctrl)[, 1])
# fit <- lm(my_otu_y ~ week)
# if(summary(fit)$r.squared > .5) {index <- 1
# } else {index <- 0}
# index
# 
# rsq <- NULL
# for (i in 1:10) {
#   my_otu_y <- as.vector(otu_table(ps_LF_ctrl)[, i])
#   rsq[i] <- summary(lm(my_otu_y ~ week))$r.squared
# }
# rsq

# check out the fourth one
# my_otu_y <- as.vector(otu_table(ps_LF_ctrl)[, 4])
# fit <- lm(my_otu_y ~ week)
# summary(fit)
# 
# my_otu_y
```

```{r}
# define x variable: time (weeks)
# week <- c(1, 12, 2:7)
# index <- rep(0, dim(otu_table(ps_LF_ctrl))[2])
# for(i in 1:dim(otu_table(ps_LF_ctrl))[2]){
#   # define y variable: relative abundance
#   my_otu_y <- as.vector(otu_table(ps_LF_ctrl)[, i])
#   fit <- lm(my_otu_y ~ week)
#   summary <- summary(fit)
#   if(summary(fit)$r.squared > .5) index[i] <- 1
# }
# sum(index)
```


